---
title: "MWAS analysis"
author: "Hongxu Wang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{css, echo=FALSE}
pre {
  max-height: 200px;
  overflow-: auto;
}
```


```{r mwas data setup, include = FALSE, echo = FALSE}
source(fs::path(here::here(),"0_project_setup","!libraries.R"))
source(fs::path(here::here(),"0_project_setup","!directories.R"))
source(fs::path(here::here(),"0_project_setup","!set_exposure_outcome_vars.R"))

sol_df <- read_rds(fs::path(dir_data_analysis, "SOLAR analysis full data.rds"))

solar_ft <- read_rds(fs::path(dir_data_analysis, "SOLAR feature data all modes.rds"))

options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE)
```

## MWAS Analysis overview

**HEADER**: metabolomics - outcome analysis\
**NAME OF analysis dataset**: sol_df, chs_df\
**Type of model**: 
glm(diabete_change ~ metabolites + covars(sex, bmi, age, hispanic(CHS only),fasting glucose, total cholesterol, triglycerides))\
**Function/package used to fit the model**: epiomics::owas()
**EXPOSURES**:
Metabolomics feature data\
**OUTCOMES**:
diabete_change\
**Covariates**:
**solar covars**: sex, bmi, age, fasting glucose, total cholesterol, triglycerides (for diabete_change outcome)\
**chs covars**:sex, bmi, age, hispanic,fasting glucose, total cholesterol, triglycerides\

## MWAS analysis
```{r mwas analysis, echo = TRUE}
# variables of outcome, covars, metabolites-----
outcomes <- c("diabete_change_di")

covars <- c("age_1", 
            "sex_1", 
            "bmi_1", 
            "hisp_1",
            "og_glu_5_1", 
            "og_glu120_1",
            "tot_chol_1",
            "tag_1")

metabolites <-
  colnames(solar_ft)[2:ncol(solar_ft)]

# Solar OWAS function----
sol_df_mwas <- sol_df %>%
    mutate(diabete_change_di = ifelse(diabete_change_di == "Yes", 1, 0))

mwas_solar <- epiomics::owas(sol_df_mwas,
                var = outcomes,
                omics = metabolites,
                covars = covars[-4],
                var_exposure_or_outcome = "outcome",
                family = "binomial",
                conf_int = TRUE)

# saveRDS(mwas_solar,
#         file = fs::path(dir_results,
#                         "1_1 SOLAR mwas results OWAS.rds"))


# (1) merge with feature metadata to include the mode variable
# (2) add significance variable
mwas_solar <- read_rds(fs::path(dir_results,
                        "1_1 SOLAR mwas results OWAS.rds"))

# Load Metabolomics Feature Tables 
ftdata <- read_rds(fs::path(dir_data,
                   "sol_chs_batch_cor_scaled_untargeted_fts.rds"))

# Obtain Feature metadata 
ft_metadata <- ftdata$solar %>% 
  modify(~data.frame(feature = colnames(.)[-1])) %>% 
  bind_rows(.id = "mode")

mwas_solar1 <- mwas_solar %>% tidylog::left_join(ft_metadata, by = c("feature_name" = "feature")) %>% 
  mutate(sig = ifelse(p_value < 0.05, "Sig.", "Not Sig."),
         mode = ifelse(str_detect(mode, "neg"), "negative", "positive")) 


# separate mz/rt variables
mwas_solar2 <- mwas_solar1  %>% mutate(
                 m.z = str_split_fixed(feature_name, "_", n = 2)[,1] %>%
                   as.numeric(.),
                 r.t = str_split_fixed(feature_name, "_", n = 2)[,2] %>%
                   as.numeric(.))

```


## MWAS Analysis Summary of P value
```{r mwas analysis summary, echo = TRUE}
# summary table
(solar_summary_mwas <- mwas_solar2 %>% 
  group_by(var_name) %>% 
  summarise(n_features = length(feature_name), 
            percent_significant_p05 = npct(sig, 
                                           "Sig.", 
                                           n.digits = 2))) #1105

# rm(mwas_solar, mwas_solar1, solar_data, 
#     sol_df, sol_df_mwas, solar_ft, solar_summary_mwas, )
```



