---
title: "Mummichog pathway analysis"
author: "Hongxu Wang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{css, echo=FALSE}
pre {
  max-height: 200px;
  overflow-: auto;
}
```


```{r pathway analysis data setup, include = FALSE, echo = FALSE}
source(fs::path(here::here(),"0_project_setup","!libraries.R"))
source(fs::path(here::here(),"0_project_setup","!directories.R"))
source(fs::path(here::here(),"0_project_setup","!set_exposure_outcome_vars.R"))

options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE)
```

## Set up Dataset for mummichog
```{r set up Dataset for mummichog, echo = TRUE, warning = FALSE}
# Set up folder structure------- 
for(outcome_name in outcomes){
  # Create outcome Folder: Level 1 (outcome)
  dir.create(file.path(fs::path(dir_results, 
                                outcome_name)), 
             showWarnings = TRUE)
  # Level 2 (Cohort): Create SOLAR Folder
  dir.create(file.path(fs::path(dir_results, 
                                outcome_name, 
                                "solar")), 
             showWarnings = TRUE)
  rm(outcome_name)
}

# Loop over outcomes, cohort to save mwas
for(item in outcomes){
    temp_mwas <- mwas_solar2 %>%
      filter(var_name == item) %>%
      dplyr::rename(p.value = p_value, 
                    t.score = test_statistic) %>%
      dplyr::select(m.z, p.value, t.score, r.t, mode)
      write_csv(temp_mwas, 
                fs::path(dir_results, item,
                         "solar",
                         paste("solar", 
                               item, 
                               "MWAS.csv",
                               sep = "_")))
  }
```

## Run Mummichog

To run mummichog analysis, use the data saved above

Then, run mummichog on the metaboanalyst website (https://www.metaboanalyst.ca/)

Click "start here" --> "functional analysis" you should be prompted to upload a peak list file

These next steps are important!!!

select:

Ion mode: Mixed mode

Mass tolerance 5.0

Retention time: Yes- seconds

Ranked by (1 column only): t-scores

Enforce Primary Ions (V2 only): should be selected

Click submit

On the next page, you will be prompted to Specify analysis parameters:

under Algorithms, select BOTH Mummichog AND GSEA

p-value cutoff: 0.05

Version 2.0

under "Select a pathway library:"

select Homo sapiens (human) [MFN] This (should be the defualt)

At the bottom of the page, select only pathways containing at least 3 entries

After you get the results, click "download" on the upper left. Download the zip file with all results. Unzip file, rename the folder as appropriate, and move to the Results to each corresponding exposures folders here: fs::path(dir_results)


