---
title: "Data Description"
author: "Hongxu Wang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r data setup, include = FALSE, echo = FALSE}

source(fs::path(here::here(),"0_project_setup","!libraries.R"))
source(fs::path(here::here(),"0_project_setup","!directories.R"))
source(fs::path(here::here(),"0_project_setup","!set_exposure_outcome_vars.R"))

options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## Data Loading 
Two cleaned datasets were loaded in this project.
One is Metabolomics Feature data, another is Exposure Outcome Data.

```{r data loading, echo = TRUE}
# Load Metabolomics Feature Tables 
ftdata <- read_rds(fs::path(dir_data,
                   "sol_chs_batch_cor_scaled_untargeted_fts.rds"))

# Obtain Feature metadata 
ft_metadata <- ftdata$solar %>% 
  modify(~data.frame(feature = colnames(.)[-1])) %>% 
  bind_rows(.id = "mode")

# Load Exposure Outcome Data from drive  ------------------------
sol <- read_rds(fs::path(dir_data,
                         "SOLAR exposure outcome data HRE PFAS.rds"))

chs <- read_rds(fs::path(dir_data_analysis,
                                  "CHS MetaAir MetaChem Cleaned Redcap and Exposures Outcome Data long V2.rds"))
```

## Data Preprocessing 
Preprocess outcome and covariates data for solar cohort
```{r data preprocessing solar, echo = TRUE, message = FALSE}
# Step1: create prediabete and diabete varaibles
# Step2: create diabete_change variable which indicate how sample's diabete status(prediabete, diabete, no diabete) change from first visit to second visit.
# Step3: create diabete_change_di variable which has two categories: no to no (NO), no to yes(Yes). 
solar_df <- sol$longitudinal %>%
  dplyr::select(id, visit, og_glu_5, og_glu120, age, sex, bmi, tot_chol, tag, a1c) %>%
  mutate(
    prediabete = case_when(
    a1c <= 6.4 & a1c >= 5.7 ~ "Yes", 
    og_glu_5 >= 100 & og_glu_5 <= 125 ~ "Yes",
    og_glu120 >= 140 & og_glu120 <= 199 ~ "Yes", 
    is.na(a1c) & is.na(og_glu_5) & is.na(og_glu120) ~ NA_character_,
    TRUE ~ "No"
    ), 
         diabete = case_when(
           a1c > 6.4 ~ "Yes",
           og_glu_5 > 125 ~ "Yes",
           og_glu120 > 199 ~ "Yes",
           is.na(a1c) & is.na(og_glu_5) & is.na(og_glu120) ~
             NA_character_,
           TRUE ~ "No"
         ),
    prediabete_or_diabete = case_when(
      prediabete == "Yes"|diabete == "Yes" ~ "Yes",
      prediabete == "No" & diabete == "No" ~ "No"
    )
  ) %>%
  mutate(prediabete_or_diabete = as.factor(prediabete_or_diabete))

# pivot wider
solar_df_w <- solar_df %>% 
  pivot_wider(
    names_from = "visit",
    values_from = og_glu_5:prediabete_or_diabete
  )

solar_df_w1 <- solar_df_w %>%
  mutate(
         diabete_change = case_when(
           prediabete_or_diabete_1 == "No"&
             prediabete_or_diabete_2 == "No" ~ "No to no",
           prediabete_1 == "Yes" & 
             prediabete_2 == "Yes" ~ "Prediabete to prediabete",
           diabete_1 == "Yes" &
             diabete_2 == "Yes" ~ "Diabete to diabete",
           prediabete_or_diabete_1 == "No" &
             prediabete_2 == "Yes" ~ "No to prediabete",
           prediabete_or_diabete_1 == "No" &
             diabete_2 == "Yes" ~ "No to diabete",
           prediabete_1 == "Yes" &
             diabete_2 == "Yes" ~ "Prediabete to diabete",
           prediabete_1 =="Yes" &
             prediabete_or_diabete_2 == "No" ~ "Prediabete to no",
           diabete_1 == "Yes" & 
             prediabete_2 == "Yes" ~ "Diabete to prediabete",
           diabete_1 == "Yes" &
             prediabete_or_diabete_2 == "No" ~ "Diabete to no"))

## Adding missing participants to the data. 
missing_solar_df_w<- solar_df_w %>%
  filter(id == "r02_sol_00078"|id == "r02_sol_00178") %>%
  mutate(
         diabete_change = case_when(
           prediabete_or_diabete_2 == "No"&
             prediabete_or_diabete_3 == "No" ~ "No to no",
           prediabete_2 == "Yes" & 
             prediabete_3 == "Yes" ~ "Prediabete to prediabete",
           diabete_2 == "Yes" &
             diabete_3 == "Yes" ~ "Diabete to diabete",
           prediabete_or_diabete_2 == "No" &
             prediabete_3 == "Yes" ~ "No to prediabete",
           prediabete_or_diabete_2 == "No" &
             diabete_3 == "Yes" ~ "No to diabete",
           prediabete_2 == "Yes" &
             diabete_3 == "Yes" ~ "Prediabete to diabete",
           prediabete_2 =="Yes" &
             prediabete_or_diabete_3 == "No" ~ "Prediabete to no",
           diabete_2 == "Yes" & 
             prediabete_3 == "Yes" ~ "Diabete to prediabete",
           diabete_2 == "Yes" &
             prediabete_or_diabete_3 == "No" ~ "Diabete to no"))

missing_solar_df_w2<- solar_df_w %>%
  filter(!is.na(og_glu_5_1) & is.na(og_glu_5_2) & !is.na(og_glu_5_3)) %>%
  mutate(
         diabete_change = case_when(
           prediabete_or_diabete_1 == "No"&
             prediabete_or_diabete_3 == "No" ~ "No to no",
           prediabete_1 == "Yes" & 
             prediabete_3 == "Yes" ~ "Prediabete to prediabete",
           diabete_1 == "Yes" &
             diabete_3 == "Yes" ~ "Diabete to diabete",
           prediabete_or_diabete_1 == "No" &
             prediabete_3 == "Yes" ~ "No to prediabete",
           prediabete_or_diabete_1 == "No" &
             diabete_3 == "Yes" ~ "No to diabete",
           prediabete_1 == "Yes" &
             diabete_3 == "Yes" ~ "Prediabete to diabete",
           prediabete_1 =="Yes" &
             prediabete_or_diabete_3 == "No" ~ "Prediabete to no",
           diabete_1 == "Yes" & 
             prediabete_3 == "Yes" ~ "Diabete to prediabete",
           diabete_1 == "Yes" &
             prediabete_or_diabete_3 == "No" ~ "Diabete to no"))
  
missing_solar_df_w3<- solar_df_w %>%
  filter(!is.na(og_glu_5_1) & is.na(og_glu_5_2) & 
           is.na(og_glu_5_3) & !is.na(og_glu_5_4)) %>%
  mutate(
         diabete_change = case_when(
           prediabete_or_diabete_1 == "No"&
             prediabete_or_diabete_4 == "No" ~ "No to no",
           prediabete_1 == "Yes" & 
             prediabete_4 == "Yes" ~ "Prediabete to prediabete",
           diabete_1 == "Yes" &
             diabete_4 == "Yes" ~ "Diabete to diabete",
           prediabete_or_diabete_1 == "No" &
             prediabete_4 == "Yes" ~ "No to prediabete",
           prediabete_or_diabete_1 == "No" &
             diabete_4 == "Yes" ~ "No to diabete",
           prediabete_1 == "Yes" &
             diabete_4 == "Yes" ~ "Prediabete to diabete",
           prediabete_1 =="Yes" &
             prediabete_or_diabete_4 == "No" ~ "Prediabete to no",
           diabete_1 == "Yes" & 
             prediabete_4 == "Yes" ~ "Diabete to prediabete",
           diabete_1 == "Yes" &
             prediabete_or_diabete_4 == "No" ~ "Diabete to no"))

missing_solar_df_w4<- solar_df_w %>%
  filter(!is.na(og_glu_5_1) & is.na(og_glu_5_2) & 
           is.na(og_glu_5_3) & is.na(og_glu_5_4) &
           !is.na(og_glu_5_5)) %>%
  mutate(
         diabete_change = case_when(
           prediabete_or_diabete_1 == "No"&
             prediabete_or_diabete_5 == "No" ~ "No to no",
           prediabete_1 == "Yes" & 
             prediabete_5 == "Yes" ~ "Prediabete to prediabete",
           diabete_1 == "Yes" &
             diabete_5 == "Yes" ~ "Diabete to diabete",
           prediabete_or_diabete_1 == "No" &
             prediabete_5 == "Yes" ~ "No to prediabete",
           prediabete_or_diabete_1 == "No" &
             diabete_5 == "Yes" ~ "No to diabete",
           prediabete_1 == "Yes" &
             diabete_5 == "Yes" ~ "Prediabete to diabete",
           prediabete_1 =="Yes" &
             prediabete_or_diabete_5 == "No" ~ "Prediabete to no",
           diabete_1 == "Yes" & 
             prediabete_5 == "Yes" ~ "Diabete to prediabete",
           diabete_1 == "Yes" &
             prediabete_or_diabete_5 == "No" ~ "Diabete to no"))

missing_solar_df_w5<- solar_df_w %>%
  filter(!is.na(og_glu_5_1) & is.na(og_glu_5_2) & 
           is.na(og_glu_5_3) & is.na(og_glu_5_4) &
           is.na(og_glu_5_5) & !is.na(og_glu_5_6)) %>%
    mutate(
         diabete_change = case_when(
           prediabete_or_diabete_1 == "No"&
             prediabete_or_diabete_6 == "No" ~ "No to no",
           prediabete_1 == "Yes" & 
             prediabete_6 == "Yes" ~ "Prediabete to prediabete",
           diabete_1 == "Yes" &
             diabete_6 == "Yes" ~ "Diabete to diabete",
           prediabete_or_diabete_1 == "No" &
             prediabete_6 == "Yes" ~ "No to prediabete",
           prediabete_or_diabete_1 == "No" &
             diabete_6 == "Yes" ~ "No to diabete",
           prediabete_1 == "Yes" &
             diabete_6 == "Yes" ~ "Prediabete to diabete",
           prediabete_1 =="Yes" &
             prediabete_or_diabete_6 == "No" ~ "Prediabete to no",
           diabete_1 == "Yes" & 
             prediabete_6 == "Yes" ~ "Diabete to prediabete",
           diabete_1 == "Yes" &
             prediabete_or_diabete_6 == "No" ~ "Diabete to no"))

missing_solar_df_w6<- solar_df_w %>%
  filter(!is.na(og_glu_5_1) & is.na(og_glu_5_2) & 
           is.na(og_glu_5_3) & is.na(og_glu_5_4) &
           is.na(og_glu_5_5) & is.na(og_glu_5_6) &
           !is.na(og_glu_5_7)) %>%
    mutate(
         diabete_change = case_when(
           prediabete_or_diabete_1 == "No"&
             prediabete_or_diabete_7 == "No" ~ "No to no",
           prediabete_1 == "Yes" & 
             prediabete_7 == "Yes" ~ "Prediabete to prediabete",
           diabete_1 == "Yes" &
             diabete_7 == "Yes" ~ "Diabete to diabete",
           prediabete_or_diabete_1 == "No" &
             prediabete_7 == "Yes" ~ "No to prediabete",
           prediabete_or_diabete_1 == "No" &
             diabete_7 == "Yes" ~ "No to diabete",
           prediabete_1 == "Yes" &
             diabete_7 == "Yes" ~ "Prediabete to diabete",
           prediabete_1 =="Yes" &
             prediabete_or_diabete_7 == "No" ~ "Prediabete to no",
           diabete_1 == "Yes" & 
             prediabete_7 == "Yes" ~ "Diabete to prediabete",
           diabete_1 == "Yes" &
             prediabete_or_diabete_7 == "No" ~ "Diabete to no"))

solar_df_w1 <- solar_df_w1 %>% 
  filter(!id %in% missing_solar_df_w$id & 
           !id %in% missing_solar_df_w2$id & 
           !id %in% missing_solar_df_w3$id &
           !id %in% missing_solar_df_w4$id &
           !id %in% missing_solar_df_w5$id &
           !id %in% missing_solar_df_w6$id) %>% 
  bind_rows(missing_solar_df_w) %>%
  bind_rows(missing_solar_df_w2) %>%
  bind_rows(missing_solar_df_w3) %>%
  bind_rows(missing_solar_df_w4) %>%
  bind_rows(missing_solar_df_w5) %>%
  bind_rows(missing_solar_df_w6) 

table1::table1(~diabete_change, solar_df_w1)

solar_df_w2<- solar_df_w1 %>% mutate(
  diabete_change_di = case_when(
           diabete_change == "No to diabete" ~ "No to Yes",
           diabete_change == "No to prediabete" ~ "No to Yes",
           TRUE ~ diabete_change
         ),
  diabete_change_di = factor(diabete_change_di))
```

Preprocess outcome and covars data for chs cohort
```{r data preprocessing chs, echo = TRUE, message = FALSE}
chs_df <- chs %>%
  dplyr::select(id, study, og_glu_5, og_glu120, ma_age, ma_sex, ma_bmi, ma_hisp, total_chol_dori, triglycerides_dori, hba1c) %>%
  rename(age = ma_age, 
         sex = ma_sex,
         bmi = ma_bmi,
         hisp = ma_hisp,
         tot_chol = total_chol_dori, 
         tag = triglycerides_dori,
         a1c = hba1c) %>%
  mutate(
    prediabete = case_when(
    a1c <= 6.4 & a1c >= 5.7 ~ "Yes", 
    og_glu_5 >= 100 & og_glu_5 <= 125 ~ "Yes",
    og_glu120 >= 140 & og_glu120 <= 199 ~ "Yes", 
    is.na(a1c) & is.na(og_glu_5) & is.na(og_glu120) ~ NA_character_,
    TRUE ~ "No"
    ), 
         diabete = case_when(
           a1c > 6.4 ~ "Yes",
           og_glu_5 > 125 ~ "Yes",
           og_glu120 > 199 ~ "Yes",
           is.na(a1c) & is.na(og_glu_5) & is.na(og_glu120) ~ NA_character_,
           TRUE ~ "No"
         ),
    prediabete_or_diabete = case_when(
      prediabete == "Yes"|diabete == "Yes" ~ "Yes",
      prediabete == "No" & diabete == "No" ~ "No"
    )
  ) %>%
  mutate(prediabete_or_diabete = as.factor(prediabete_or_diabete),
         visit = ifelse(study == "MetaAir", 1, 2)) %>%
  select(id, visit, age, everything())

# pivot wider
chs_df_w <- chs_df %>% 
  pivot_wider(
    names_from = "visit",
    values_from = age:prediabete_or_diabete
  )

chs_df_w1 <- chs_df_w %>%
  mutate(
         diabete_change = case_when(
           prediabete_or_diabete_1 == "No"&
             prediabete_or_diabete_2 == "No" ~ "No to no",
           prediabete_1 == "Yes" & 
             prediabete_2 == "Yes" ~ "Prediabete to prediabete",
           diabete_1 == "Yes" &
             diabete_2 == "Yes" ~ "Diabete to diabete",
           prediabete_or_diabete_1 == "No" &
             prediabete_2 == "Yes" ~ "No to prediabete",
           prediabete_or_diabete_1 == "No" &
             diabete_2 == "Yes" ~ "No to diabete",
           prediabete_1 == "Yes" &
             diabete_2 == "Yes" ~ "Prediabete to diabete",
           prediabete_1 =="Yes" &
             prediabete_or_diabete_2 == "No" ~ "Prediabete to no",
           diabete_1 == "Yes" & 
             prediabete_2 == "Yes" ~ "Diabete to prediabete",
           diabete_1 == "Yes" &
             prediabete_or_diabete_2 == "No" ~ "Diabete to no"
         ))
table1::table1(~diabete_change, chs_df_w1)
chs_df_w2 <- chs_df_w1 %>% 
  mutate(diabete_change_di = case_when(
           diabete_change == "No to prediabete" ~ "No to Yes",
           TRUE ~ diabete_change
         ),
         diabete_change_di = factor(diabete_change_di))

# Saving analysis ready dataset------
# write_rds(chs_df_w2, fs::path(dir_data_analysis, "CHS exposure outcome data HRE PFAS wide.rds"))
# write_rds(solar_df_w2, fs::path(dir_data_analysis, "SOLAR exposure outcome data HRE PFAS wide.rds"))
```

Analysis ready feature data
```{r feature data, echo = TRUE, message = FALSE}
# Preparing analysis ready feature data------
# combine four different mode.
## solar------
feature_solar <- reduce(ftdata$sol, left_join) %>%
  mutate(id = as.character(id))

## checking the number of features been imputed
# feature_solar1<- feature_solar %>%
#   mutate_at(.vars = vars(colnames(feature_solar)[-1]),
#             .funs = list(imputed = ~ifelse(abs(.)>abs(mean(.))+ 7*sd(.), 1, 0)))
# 
# feature_solar_l <- feature_solar1 %>% select(contains("imputed")) %>%
#   pivot_longer(cols = colnames(feature_solar1 %>% select(contains("imputed"))),
#                names_to = "feature",
#                values_to = "is_imputed") %>%
#   group_by(feature) %>%
#   summarise(num_of_imputation = sum(is_imputed)) %>%
#   ungroup()
# 
# feature_solar_imputed_feature_df <- feature_solar_l %>%
#   filter(num_of_imputation >0) 
#note: number of features been imputed is 5815. 


feature_solar_imputed <- feature_solar %>%
  mutate_at(.vars = vars(colnames(feature_solar)[-1]),
            .funs = ~ifelse(abs(.) > abs(mean(.))+ 7*sd(.), mean(.), .))


## chs-------
feature_chs <- reduce(ftdata$chs, left_join) %>%
  mutate(id = as.character(id))

## checking the number of features been imputed
# feature_chs1<- feature_chs %>%
#   mutate_at(.vars = vars(colnames(feature_chs)[-1]),
#             .funs = list(imputed = ~ifelse(abs(.)>abs(mean(.))+ 7*sd(.), 1, 0)))
# 
# feature_chs_l <- feature_chs1 %>% select(contains("imputed")) %>%
#   pivot_longer(cols = colnames(feature_chs1 %>% select(contains("imputed"))),
#                names_to = "feature",
#                values_to = "is_imputed") %>%
#   group_by(feature) %>%
#   summarise(num_of_imputation = sum(is_imputed)) %>%
#   ungroup()
# 
# feature_chs_imputed_feature_df <- feature_chs_l %>%
#   filter(num_of_imputation >0) #2066

feature_chs_imputed <- feature_chs %>%
  mutate_at(.vars = vars(colnames(feature_chs)[-1]),
            .funs = ~ifelse(abs(.) > abs(mean(.))+ 7*sd(.), mean(.), .))

# Saving analysis ready feature data------
# write_rds(feature_solar_imputed, fs::path(dir_data_analysis, "SOLAR feature data all modes.rds"))
# 
# write_rds(feature_chs_imputed, fs::path(dir_data_analysis, "CHS feature data all modes.rds"))

rm(solar_df, solar_df_w, solar_df_w1, 
   sol, chs, chs_df, chs_df_w,
   chs_df_w1, feature_chs,
   # feature_solar, 
   ftdata, missing_solar_df_w, missing_solar_df_w2,
   missing_solar_df_w3, missing_solar_df_w4, missing_solar_df_w5,
   missing_solar_df_w6)
```

## Merge exposure outcome data with feature data
```{r}
sol_df <- solar_df_w2 %>% tidylog::left_join(feature_solar_imputed, by = "id")
chs_df <- chs_df_w2 %>% tidylog::left_join(feature_chs_imputed, by = "id")
```


# Exclude samples with no biobanked fasting plasma, no follow-up visit, prediabetes at baseline and missing total cholesterol and triglyceride
All continuous features and covariates variables are normalized.
```{r prepare analysis ready data, echo=TRUE}
# solar -------
solar_data <-  sol_df %>% 
    mutate(across(everything(), as.vector))%>% 
    filter(diabete_change_di == "No to Yes"|diabete_change_di == "No to no") %>% # include this if outcome is diabete_change_di 
    mutate(diabete_change_di = ifelse(diabete_change_di == "No to Yes", "Yes", "No")) %>%
    drop_na(outcomes,
            all_of(covars1[-4]),
            all_of(covars2))

print(paste("SOLAR sample size is", dim(solar_data)[1]))

# chs ----------------
chs_data <- chs_df %>% 
  mutate(across(everything(), as.vector)) %>% 
  filter(prediabete_or_diabete_1 == "No") %>% # include this if outcome is diabete_change_di 
  mutate(diabete_change_di = ifelse(diabete_change_di == "No to Yes", "Yes", "No"),
         diabete_change_di = factor(diabete_change_di),
         hisp_1 = factor(hisp_1),
         sex_1 = ifelse(sex_1 == "female", "Female", "Male")) %>%
  drop_na(
          all_of(colnames(feature_chs_imputed)[-1]),
          outcomes,
          all_of(covars1),
          all_of(covars2),
          diabete_change_di)

print(paste("CHS sample size is", dim(chs_data)[1]))

write_rds(solar_data, fs::path(dir_data_analysis, "SOLAR analysis full data.rds"))

write_rds(chs_data, fs::path(dir_data_analysis, "CHS analysis full data.rds"))
```


