---
title: "Data Description"
author: "Hongxu Wang"
date: "3/23/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r data setup, include = FALSE, echo = FALSE}

source(fs::path(here::here(),"0_project_setup","!libraries.R"))
source(fs::path(here::here(),"0_project_setup","!directories.R"))

options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE)
```

## Data Loading 
Two cleaned datasets were loaded in this project.
One is Metabolomics Feature data, another is Exposure Outcome Data.

```{r data loading, echo = TRUE}
# Load Metabolomics Feature Tables 
ftdata <- read_rds(fs::path(dir_data,
                   "sol_chs_batch_cor_scaled_untargeted_fts.rds"))

# Obtain Feature metadata 
ft_metadata <- ftdata$solar %>% 
  modify(~data.frame(feature = colnames(.)[-1])) %>% 
  bind_rows(.id = "mode")

# Load Exposure Outcome Data from drive  ------------------------
sol <- read_rds(fs::path(dir_data,
                         "SOLAR exposure outcome data HRE PFAS.rds"))

chs <- read_rds(fs::path(dir_data,
                         "CHS MetaAir exposure outcome data HRE PFAS.rds"))

chs_l <- read_rds(fs::path(dir_data_chs_l,
                           "metachem_redcap_temp.rds"))

chs <- read_rds(fs::path(dir_chs, 
                                  "CHS MetaAir MetaChem Cleaned Redcap and Exposures Outcome Data wide.rds"))

# chs_analysis_l <- read_rds(fs::path(dir_chs, 
#                                   "CHS MetaAir MetaChem Cleaned Redcap and Exposures Outcome Data long.rds"))
```

## Data Preprocessing 
Adding 6 variables to solar and chs longitudinal data:

**1. og_glu_5_change**: change of Fasting (-5min) glucose from OGTT(Oral glucose tolerance test)-research from visit 1 to visit 2 (totally 12 visit).\
**2. og_glu_5_change_by_year**: og_glu_5_change/(age(visit2) - age(visit1))
**3. og_glu120_change**: change of 2-hr glucose from OGTT-research from visit 1 to visit 2.\
**4. og_glu120_change_by_year**: og_glu120_change/(age(visit2) - age(visit1))
**5. a1c_change**: change of hba1c from visit 1 to visit 2.\
**6. a1c_change_by_year**: hba1c_change/(age(visit2) - age(visit1))

Analysis ready outcome and covars data
```{r data preprocessing, echo = TRUE, message = FALSE}
# Preparing analysis ready outcome and covars dataset-----
## solar--------
solar_df <- sol$longitudinal %>%
  dplyr::select(id, visit, og_glu_5, og_glu120, age, sex, bmi, tot_chol, tot_lipid, tanner, ses, wave, a1c) %>%
  mutate(
    prediabete = case_when(
    a1c <= 6.4 & a1c >= 5.7 ~ "Yes", 
    og_glu_5 >= 100 & og_glu_5 <= 125 ~ "Yes",
    og_glu120 >= 140 & og_glu120 <= 199 ~ "Yes", 
    is.na(a1c) & is.na(og_glu_5) & is.na(og_glu120) ~ NA_character_,
    TRUE ~ "No"
    ), 
         diabete = case_when(
           a1c > 6.4 ~ "Yes",
           og_glu_5 > 125 ~ "Yes",
           og_glu120 > 199 ~ "Yes",
           is.na(a1c) & is.na(og_glu_5) & is.na(og_glu120) ~
             NA_character_,
           TRUE ~ "No"
         ),
    prediabete_or_diabete = case_when(
      prediabete == "Yes"|diabete == "Yes" ~ "Yes",
      prediabete == "No" & diabete == "No" ~ "No"
    )
  ) %>%
  mutate(prediabete_or_diabete = as.factor(prediabete_or_diabete))

# pivot wider
solar_df_w <- solar_df %>% 
  pivot_wider(
    names_from = "visit",
    values_from = og_glu_5:prediabete_or_diabete
  )
# change of fasting glucose and 2-hour glucose,
solar_df_w1 <- solar_df_w %>%
  mutate(
         age_change = age_2 - age_1,
         og_glu_5_change = og_glu_5_2 - og_glu_5_1,
         og_glu_5_change_by_year = og_glu_5_change/(age_2 - age_1),
         og_glu120_change = og_glu120_2 - og_glu120_1,
         og_glu120_change_by_year = og_glu120_change/(age_2 - age_1),
         a1c_change_by_year = (a1c_2 - a1c_1)/(age_2 - age_1),
         diabete_change = case_when(
           prediabete_or_diabete_1 == "No"&
             prediabete_or_diabete_2 == "No" ~ "No to no",
           prediabete_1 == "Yes" & 
             prediabete_2 == "Yes" ~ "Prediabete to prediabete",
           diabete_1 == "Yes" &
             diabete_2 == "Yes" ~ "Diabete to diabete",
           prediabete_or_diabete_1 == "No" &
             prediabete_2 == "Yes" ~ "No to prediabete",
           prediabete_or_diabete_1 == "No" &
             diabete_2 == "Yes" ~ "No to diabete",
           prediabete_1 == "Yes" &
             diabete_2 == "Yes" ~ "Prediabete to diabete",
           prediabete_1 =="Yes" &
             prediabete_or_diabete_2 == "No" ~ "Prediabete to no",
           diabete_1 == "Yes" & 
             prediabete_2 == "Yes" ~ "Diabete to prediabete",
           diabete_1 == "Yes" &
             prediabete_or_diabete_2 == "No" ~ "Diabete to no"))

## Adding two missing participants to the data. 
missing_solar_df_w<- solar_df_w %>%
  filter(id == "r02_sol_00078"|id == "r02_sol_00178") %>%
  mutate(
         age_change = age_3 - age_2,
         og_glu_5_change = og_glu_5_3 - og_glu_5_2,
         og_glu_5_change_by_year = og_glu_5_change/(age_3 - age_2),
         og_glu120_change = og_glu120_3 - og_glu120_2,
         og_glu120_change_by_year = og_glu120_change/(age_3 - age_2),
         a1c_change_by_year = (a1c_3 - a1c_2)/(age_3 - age_2),
         diabete_change = case_when(
           prediabete_or_diabete_2 == "No"&
             prediabete_or_diabete_3 == "No" ~ "No to no",
           prediabete_2 == "Yes" & 
             prediabete_3 == "Yes" ~ "Prediabete to prediabete",
           diabete_2 == "Yes" &
             diabete_3 == "Yes" ~ "Diabete to diabete",
           prediabete_or_diabete_2 == "No" &
             prediabete_3 == "Yes" ~ "No to prediabete",
           prediabete_or_diabete_2 == "No" &
             diabete_3 == "Yes" ~ "No to diabete",
           prediabete_2 == "Yes" &
             diabete_3 == "Yes" ~ "Prediabete to diabete",
           prediabete_2 =="Yes" &
             prediabete_or_diabete_3 == "No" ~ "Prediabete to no",
           diabete_2 == "Yes" & 
             prediabete_3 == "Yes" ~ "Diabete to prediabete",
           diabete_2 == "Yes" &
             prediabete_or_diabete_3 == "No" ~ "Diabete to no"))

missing_solar_df_w2<- solar_df_w %>%
  filter(!is.na(og_glu_5_1) & is.na(og_glu_5_2) & !is.na(og_glu_5_3)) %>%
  mutate(
         age_change = age_3 - age_1,
         og_glu_5_change = og_glu_5_3 - og_glu_5_1,
         og_glu_5_change_by_year = og_glu_5_change/(age_3 - age_1),
         og_glu120_change = og_glu120_3 - og_glu120_1,
         og_glu120_change_by_year = og_glu120_change/(age_3 - age_1),
         a1c_change_by_year = (a1c_3 - a1c_1)/(age_3 - age_1),
         diabete_change = case_when(
           prediabete_or_diabete_1 == "No"&
             prediabete_or_diabete_3 == "No" ~ "No to no",
           prediabete_1 == "Yes" & 
             prediabete_3 == "Yes" ~ "Prediabete to prediabete",
           diabete_1 == "Yes" &
             diabete_3 == "Yes" ~ "Diabete to diabete",
           prediabete_or_diabete_1 == "No" &
             prediabete_3 == "Yes" ~ "No to prediabete",
           prediabete_or_diabete_1 == "No" &
             diabete_3 == "Yes" ~ "No to diabete",
           prediabete_1 == "Yes" &
             diabete_3 == "Yes" ~ "Prediabete to diabete",
           prediabete_1 =="Yes" &
             prediabete_or_diabete_3 == "No" ~ "Prediabete to no",
           diabete_1 == "Yes" & 
             prediabete_3 == "Yes" ~ "Diabete to prediabete",
           diabete_1 == "Yes" &
             prediabete_or_diabete_3 == "No" ~ "Diabete to no"))
  
missing_solar_df_w3<- solar_df_w %>%
  filter(!is.na(og_glu_5_1) & is.na(og_glu_5_2) & 
           is.na(og_glu_5_3) & !is.na(og_glu_5_4)) %>%
  mutate(
         age_change = age_4 - age_1,
         og_glu_5_change = og_glu_5_4 - og_glu_5_1,
         og_glu_5_change_by_year = og_glu_5_change/(age_4 - age_1),
         og_glu120_change = og_glu120_4 - og_glu120_1,
         og_glu120_change_by_year = og_glu120_change/(age_4 - age_1),
         a1c_change_by_year = (a1c_4 - a1c_1)/(age_4 - age_1),
         diabete_change = case_when(
           prediabete_or_diabete_1 == "No"&
             prediabete_or_diabete_4 == "No" ~ "No to no",
           prediabete_1 == "Yes" & 
             prediabete_4 == "Yes" ~ "Prediabete to prediabete",
           diabete_1 == "Yes" &
             diabete_4 == "Yes" ~ "Diabete to diabete",
           prediabete_or_diabete_1 == "No" &
             prediabete_4 == "Yes" ~ "No to prediabete",
           prediabete_or_diabete_1 == "No" &
             diabete_4 == "Yes" ~ "No to diabete",
           prediabete_1 == "Yes" &
             diabete_4 == "Yes" ~ "Prediabete to diabete",
           prediabete_1 =="Yes" &
             prediabete_or_diabete_4 == "No" ~ "Prediabete to no",
           diabete_1 == "Yes" & 
             prediabete_4 == "Yes" ~ "Diabete to prediabete",
           diabete_1 == "Yes" &
             prediabete_or_diabete_4 == "No" ~ "Diabete to no"))

missing_solar_df_w4<- solar_df_w %>%
  filter(!is.na(og_glu_5_1) & is.na(og_glu_5_2) & 
           is.na(og_glu_5_3) & is.na(og_glu_5_4) &
           !is.na(og_glu_5_5)) %>%
  mutate(
         age_change = age_5 - age_1,
         og_glu_5_change = og_glu_5_5 - og_glu_5_1,
         og_glu_5_change_by_year = og_glu_5_change/(age_5 - age_1),
         og_glu120_change = og_glu120_5 - og_glu120_1,
         og_glu120_change_by_year = og_glu120_change/(age_5 - age_1),
         a1c_change_by_year = (a1c_5 - a1c_1)/(age_5 - age_1),
         diabete_change = case_when(
           prediabete_or_diabete_1 == "No"&
             prediabete_or_diabete_5 == "No" ~ "No to no",
           prediabete_1 == "Yes" & 
             prediabete_5 == "Yes" ~ "Prediabete to prediabete",
           diabete_1 == "Yes" &
             diabete_5 == "Yes" ~ "Diabete to diabete",
           prediabete_or_diabete_1 == "No" &
             prediabete_5 == "Yes" ~ "No to prediabete",
           prediabete_or_diabete_1 == "No" &
             diabete_5 == "Yes" ~ "No to diabete",
           prediabete_1 == "Yes" &
             diabete_5 == "Yes" ~ "Prediabete to diabete",
           prediabete_1 =="Yes" &
             prediabete_or_diabete_5 == "No" ~ "Prediabete to no",
           diabete_1 == "Yes" & 
             prediabete_5 == "Yes" ~ "Diabete to prediabete",
           diabete_1 == "Yes" &
             prediabete_or_diabete_5 == "No" ~ "Diabete to no"))

missing_solar_df_w5<- solar_df_w %>%
  filter(!is.na(og_glu_5_1) & is.na(og_glu_5_2) & 
           is.na(og_glu_5_3) & is.na(og_glu_5_4) &
           is.na(og_glu_5_5) & !is.na(og_glu_5_6)) %>%
    mutate(
         age_change = age_6 - age_1,
         og_glu_5_change = og_glu_5_6 - og_glu_5_1,
         og_glu_5_change_by_year = og_glu_5_change/(age_6 - age_1),
         og_glu120_change = og_glu120_6 - og_glu120_1,
         og_glu120_change_by_year = og_glu120_change/(age_6 - age_1),
         a1c_change_by_year = (a1c_6 - a1c_1)/(age_6 - age_1),
         diabete_change = case_when(
           prediabete_or_diabete_1 == "No"&
             prediabete_or_diabete_6 == "No" ~ "No to no",
           prediabete_1 == "Yes" & 
             prediabete_6 == "Yes" ~ "Prediabete to prediabete",
           diabete_1 == "Yes" &
             diabete_6 == "Yes" ~ "Diabete to diabete",
           prediabete_or_diabete_1 == "No" &
             prediabete_6 == "Yes" ~ "No to prediabete",
           prediabete_or_diabete_1 == "No" &
             diabete_6 == "Yes" ~ "No to diabete",
           prediabete_1 == "Yes" &
             diabete_6 == "Yes" ~ "Prediabete to diabete",
           prediabete_1 =="Yes" &
             prediabete_or_diabete_6 == "No" ~ "Prediabete to no",
           diabete_1 == "Yes" & 
             prediabete_6 == "Yes" ~ "Diabete to prediabete",
           diabete_1 == "Yes" &
             prediabete_or_diabete_6 == "No" ~ "Diabete to no"))

missing_solar_df_w6<- solar_df_w %>%
  filter(!is.na(og_glu_5_1) & is.na(og_glu_5_2) & 
           is.na(og_glu_5_3) & is.na(og_glu_5_4) &
           is.na(og_glu_5_5) & is.na(og_glu_5_6) &
           !is.na(og_glu_5_7)) %>%
    mutate(
         age_change = age_7 - age_1,
         og_glu_5_change = og_glu_5_7 - og_glu_5_1,
         og_glu_5_change_by_year = og_glu_5_change/(age_7 - age_1),
         og_glu120_change = og_glu120_7 - og_glu120_1,
         og_glu120_change_by_year = og_glu120_change/(age_7 - age_1),
         a1c_change_by_year = (a1c_7 - a1c_1)/(age_7 - age_1),
         diabete_change = case_when(
           prediabete_or_diabete_1 == "No"&
             prediabete_or_diabete_7 == "No" ~ "No to no",
           prediabete_1 == "Yes" & 
             prediabete_7 == "Yes" ~ "Prediabete to prediabete",
           diabete_1 == "Yes" &
             diabete_7 == "Yes" ~ "Diabete to diabete",
           prediabete_or_diabete_1 == "No" &
             prediabete_7 == "Yes" ~ "No to prediabete",
           prediabete_or_diabete_1 == "No" &
             diabete_7 == "Yes" ~ "No to diabete",
           prediabete_1 == "Yes" &
             diabete_7 == "Yes" ~ "Prediabete to diabete",
           prediabete_1 =="Yes" &
             prediabete_or_diabete_7 == "No" ~ "Prediabete to no",
           diabete_1 == "Yes" & 
             prediabete_7 == "Yes" ~ "Diabete to prediabete",
           diabete_1 == "Yes" &
             prediabete_or_diabete_7 == "No" ~ "Diabete to no"))

solar_df_w1 <- solar_df_w1 %>% 
  filter(!id %in% missing_solar_df_w$id & 
           !id %in% missing_solar_df_w2$id & 
           !id %in% missing_solar_df_w3$id &
           !id %in% missing_solar_df_w4$id &
           !id %in% missing_solar_df_w5$id &
           !id %in% missing_solar_df_w6$id) %>% 
  bind_rows(missing_solar_df_w) %>%
  bind_rows(missing_solar_df_w2) %>%
  bind_rows(missing_solar_df_w3) %>%
  bind_rows(missing_solar_df_w4) %>%
  bind_rows(missing_solar_df_w5) %>%
  bind_rows(missing_solar_df_w6) 

solar_df_w2<- solar_df_w1 %>% mutate(
  diabete_change_di = case_when(
           diabete_change == "No to diabete" ~ "No to Yes",
           diabete_change == "No to prediabete" ~ "No to Yes",
           TRUE ~ diabete_change
         ),
  diabete_change_di = factor(diabete_change_di))

## chs-----
chs_baseline <- chs$baseline %>%
  mutate(visit = "1",
         id = as.character(id),
         visitdate = as.Date(visitdate)) %>%
  dplyr::select(id, visit, cage, sex, bmi, race,visitdate, ses, tot_chol, tot_lipid, og_glu_5, og_glu120, a1c)
  

chs_visit2 <- chs_l %>% 
  rename(a1c = hba1c) %>%
  mutate(sex = ifelse(male_factor == "female", "Female", "Male"),
         og_glu_5 = ifelse(is.na(og_glu_5), og_glu_5_bedside, og_glu_5),
         og_glu120 = ifelse(is.na(og_glu120), og_glu120_bedside, og_glu120),
         visit = "2",
         id = as.character(id),
         # visitdate = as.Date(visitdate),
         cage = as.numeric(cage),
         bmi = as.numeric(bmi),
         a1c = as.numeric(a1c)
         ) %>%
  dplyr::select(id, visit, cage, sex,  bmi, og_glu_5, og_glu120, a1c) 

## combine baseline and followup chs data
chs_df <- tidylog::full_join(chs_baseline, chs_visit2)%>%
  mutate(
    prediabete = case_when(
    a1c <= 6.4 & a1c >= 5.7 ~ "Yes", 
    og_glu_5 >= 100 & og_glu_5 <= 125 ~ "Yes",
    og_glu120 >= 140 & og_glu120 <= 199 ~ "Yes", 
    is.na(a1c) & is.na(og_glu_5) & is.na(og_glu120) ~ NA_character_,
    TRUE ~ "No"
    ), 
         diabete = case_when(
           a1c > 6.4 ~ "Yes",
           og_glu_5 > 125 ~ "Yes",
           og_glu120 > 199 ~ "Yes",
           is.na(a1c) & is.na(og_glu_5) & is.na(og_glu120) ~ NA_character_,
           TRUE ~ "No"
         ),
    prediabete_or_diabete = case_when(
      prediabete == "Yes"|diabete == "Yes" ~ "Yes",
      prediabete == "No" & diabete == "No" ~ "No"
    )
  ) %>%
  mutate(prediabete_or_diabete = as.factor(prediabete_or_diabete),
         age = cage) %>%
  select(id, visit, age, -cage, everything())

# pivot wider
chs_df_w <- chs_df %>% 
  pivot_wider(
    names_from = "visit",
    values_from = age:prediabete_or_diabete
  )

chs_df_w1 <- chs_df_w %>%
  mutate(
         age_change = age_2 - age_1,
         og_glu_5_change = og_glu_5_2 - og_glu_5_1,
         og_glu_5_change_by_year = og_glu_5_change/(age_2 - age_1),
         og_glu120_change = og_glu120_2 - og_glu120_1,
         og_glu120_change_by_year = og_glu120_change/(age_2 - age_1),
         a1c_change_by_year = (a1c_2 - a1c_1)/(age_2 - age_1),
         diabete_change = case_when(
           prediabete_or_diabete_1 == "No"&
             prediabete_or_diabete_2 == "No" ~ "No to no",
           prediabete_1 == "Yes" & 
             prediabete_2 == "Yes" ~ "Prediabete to prediabete",
           diabete_1 == "Yes" &
             diabete_2 == "Yes" ~ "Diabete to diabete",
           prediabete_or_diabete_1 == "No" &
             prediabete_2 == "Yes" ~ "No to prediabete",
           prediabete_or_diabete_1 == "No" &
             diabete_2 == "Yes" ~ "No to diabete",
           prediabete_1 == "Yes" &
             diabete_2 == "Yes" ~ "Prediabete to diabete",
           prediabete_1 =="Yes" &
             prediabete_or_diabete_2 == "No" ~ "Prediabete to no",
           diabete_1 == "Yes" & 
             prediabete_2 == "Yes" ~ "Diabete to prediabete",
           diabete_1 == "Yes" &
             prediabete_or_diabete_2 == "No" ~ "Diabete to no"
         ))

chs_df_w2 <- chs_df_w1 %>% 
  mutate(diabete_change_di = case_when(
           diabete_change == "No to prediabete" ~ "No to Yes",
           TRUE ~ diabete_change
         ),
         diabete_change_di = factor(diabete_change_di))
# Saving analysis ready dataset------
write_rds(chs_df_w2, fs::path(dir_data_analysis, "CHS exposure outcome data HRE PFAS wide.rds"))
write_rds(solar_df_w2, fs::path(dir_data_analysis, "SOLAR exposure outcome data HRE PFAS wide.rds"))
```

Analysis ready feature data
```{r feature data, echo = TRUE, message = FALSE}
# Preparing analysis ready feature data------
# combine four different mode.
## solar------
feature_solar <- reduce(ftdata$sol, left_join) %>%
  mutate(id = as.character(id))

## chs-------
feature_chs <- reduce(ftdata$chs, left_join) %>%
  mutate(id = as.character(id))

# Saving analysis ready feature data------
write_rds(feature_solar, fs::path(dir_data_analysis, "SOLAR feature data all modes.rds"))

write_rds(feature_chs, fs::path(dir_data_analysis, "CHS feature data all modes.rds"))

rm(solar_df, solar_df_w, solar_df_w1, solar_df_w2,
   sol, chs,chs_baseline, chs_df, chs_df_w,
   chs_df_w1, chs_df_w2, chs_l, chs_visit2, feature_chs,
   feature_solar, ftdata, missing_solar_df_w, missing_solar_df_w2,
   missing_solar_df_w3, missing_solar_df_w4, missing_solar_df_w5,
   missing_solar_df_w6)
```

## Descriptive statistic
```{r descriptive statistic}
solar_stat <- solar %>% filter(diabete_change_di == "No to no"|diabete_change_di == "No to Yes")

table1::table1(~ age_change + sex_1 + bmi_1 + og_glu_5_1 + diabete_change_di, data = solar_stat )

chs_stat <- chs %>% filter(diabete_change_di == "No to no"|diabete_change_di == "No to Yes")

table1::table1(~ age_change + sex_1 + bmi_1 + og_glu_5_1 + diabete_change_di, data = chs_stat )
```

