---
title: "Feature selection"
author: "Hongxu Wang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
---


## Load data
```{r models comparison data setup, include = FALSE, echo = FALSE}
source(fs::path(here::here(),"0_project_setup","!libraries.R"))
source(fs::path(here::here(),"0_project_setup","!directories.R"))
source(fs::path(here::here(),"0_project_setup","!set_exposure_outcome_vars.R"))

sol_data <- read_rds(fs::path(dir_data_analysis, "SOLAR analysis full data.rds"))

chs_data <- read_rds(fs::path(dir_data_analysis, "CHS analysis full data.rds"))

solar_ft <- read_rds(fs::path(dir_data_analysis, "SOLAR feature data all modes.rds"))

options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE)
```


## Preparing analysis ready solar and chs data
All continuous features and covariates variables are normalized.
```{r prepare analysis ready data, echo=TRUE}
# solar -------
solar_data <- solar_data %>%
  mutate_at(.vars = vars(all_of(features),age_1, bmi_1, og_glu_5_1, og_glu120_1, tot_chol_1, tag_1 ),
            .funs = list(~scale(.))
            )

# chs ----------------
chs_data <- chs_data %>%
  mutate_at(.vars = vars(all_of(features),age_1, bmi_1, og_glu_5_1, og_glu120_1, tot_chol_1, tag_1),
            .funs = list(~scale(.)))
```


## Solar Model1:  general risk factors
```{r grf solar, echo=TRUE}
solar_data_model <- solar_data %>% 
  select(diabete_change_di,
         all_of(covars1[-4])) %>%
  mutate(diabete_change_di = factor(diabete_change_di))

fit <- glm(diabete_change_di ~.,
                 data = solar_data_model,
                 family = "binomial")

result_df1<- fit %>% tidy(., conf.level = 0.95) 

solar_pred1<- solar_data %>% 
  dplyr::select(outcomes[2]) %>%
  dplyr::mutate(pred = data.frame(pred = predict(fit,
                               type = "response"))$pred) %>%
  mutate(diabete_change_di = factor(diabete_change_di))

# auc value
# roc(solar_pred1$diabete_change_di, solar_pred1$pred)
# ci.auc(solar_pred1$diabete_change_di, solar_pred1$pred)

# vif(fit)
```

## Solar Model2: diabete_change_di ~ general risk factors + specific risk factors
```{r grf srf solar, echo=TRUE}
solar_data_model <- solar_data %>% 
  select(diabete_change_di,
         all_of(covars1[-4]),
         all_of(covars2)) %>%
  mutate(diabete_change_di = factor(diabete_change_di))

fit <- glm(diabete_change_di ~.,
                 data = solar_data_model,
                 family = "binomial")

result_df2<- fit %>% tidy(., conf.level = 0.95) 

solar_pred2<- solar_data_model %>% 
  dplyr::select(outcomes[2]) %>%
  dplyr::mutate(pred = data.frame(pred = predict(fit,
                               type = "response"))$pred) %>%
  mutate(diabete_change_di = factor(diabete_change_di))

# auc value
# roc(solar_pred2$diabete_change_di, solar_pred2$pred)
# ci.auc(solar_pred2$diabete_change_di, solar_pred2$pred)

# vif(fit)
```


## SOLAR model3: diabete_change_di ~ general risk factors + lasso-selected metabolites
```{r grf met solar, echo=TRUE}
solar_data_model <- solar_data %>% 
  select(diabete_change_di,
         all_of(selected_feature_name),
         all_of(covars1[-4])) %>%
   mutate(diabete_change_di = factor(diabete_change_di))

# logistic regression
fit <- glm(diabete_change_di~.,
           family = "binomial",
           data = solar_data_model) 

result_df3 <- fit %>% tidy(., conf.int = TRUE) %>% 
  mutate(model = "GRF and metabolites",
         odds_ratio = exp(estimate),
         exp_conf_high = exp(conf.high),
         exp_conf_low = exp(conf.low))
  
solar_pred3 <- solar_data_model %>% 
  dplyr::select(outcomes[2]) %>%
  dplyr::mutate(pred = data.frame(pred = predict(fit, type = "response"))$pred)

# roc(solar_pred3$diabete_change_di, solar_pred3$pred)
# ci.auc(solar_pred3$diabete_change_di, solar_pred3$pred)
# 
# vif(fit)

```


## SOLAR Model4: diabete_change_di ~ general risk factors + specific risk factors + lasso-selected metabolites
```{r grf srf met, echo= TRUE}
solar_data_model <- solar_data %>% 
    select(diabete_change_di, 
           all_of(selected_feature_name), 
           all_of(covars1[-4]),
           all_of(covars2))%>%
  mutate(diabete_change_di = factor(diabete_change_di))

# logistic regression
fit <- glm(diabete_change_di~.,
           family = "binomial",
           data = solar_data_model) 

result_df4 <- fit %>% tidy(conf.int = TRUE) %>%
  mutate(
         model = "GRF + SRF + metabolites",
         odds_ratio = exp(estimate),
         exp_conf_high = exp(conf.high),
         exp_conf_low = exp(conf.low))

# Adding annotation information to the result
model_result_annotation <- result_df4 %>% tidylog::left_join(mzrt_key_sub %>% mutate(feature = str_c("`", name ,"`")), by = c("term"= "feature")) %>% 
  select(-name, -pathway, -query_mass, -retention_time, -mass_diff,
         -matched_compound, -empirical_compound, -met_name, - path_2, term, met_name_first, everything())

write_csv(model_result_annotation, fs::path(dir_results,"prediction", "4_solar_model_result_with_annotation_5000.csv"))

solar_pred4 <- solar_data_model %>% 
  dplyr::select(outcomes[2]) %>%
  dplyr::mutate(pred = data.frame(pred = predict(fit, type = "response"))$pred)

# roc(solar_pred4$diabete_change_di, solar_pred4$pred)
# ci.auc(solar_pred4$diabete_change_di, solar_pred4$pred)

# vif(fit)

```


## SOLAR model5: diabete_change_di ~ selected features
```{r met solar, echo=TRUE}
solar_data_model <- solar_data %>% 
  select(diabete_change_di,
         all_of(selected_feature_name)) %>%
   mutate(diabete_change_di = factor(diabete_change_di))

# logistic regression
fit <- glm(diabete_change_di~.,
           family = "binomial",
           data = solar_data_model) 

result_df5 <- fit %>% tidy(., conf.int = TRUE) %>% 
  mutate(model = "metabolites",
         odds_ratio = exp(estimate),
         exp_conf_high = exp(conf.high),
         exp_conf_low = exp(conf.low))

solar_pred5 <- solar_data_model %>% 
  dplyr::select(outcomes[2]) %>%
  dplyr::mutate(pred = data.frame(pred = predict(fit, type = "response"))$pred)

# roc(solar_pred5$diabete_change_di, solar_pred5$pred)
# ci.auc(solar_pred5$diabete_change_di, solar_pred5$pred)
```


## CHS model1:  general risk factors 
```{r grf chs, echo=TRUE}
chs_data_model <- chs_data %>% 
  select(diabete_change_di,
         all_of(covars1)) 

fit <- glm(diabete_change_di ~.,
                 data = chs_data_model,
                 family = "binomial")

result_df1<- fit %>% tidy(., conf.int = TRUE) %>%
  mutate(model = "GRF",
         odds_ratio = exp(estimate),
         exp_conf_high = exp(conf.high),
         exp_conf_low = exp(conf.low))


chs_pred1<- chs_data %>% 
  dplyr::select(outcomes[2]) %>%
  dplyr::mutate(pred = data.frame(pred = predict(fit,
                               type = "response"))$pred) %>%
  mutate(diabete_change_di = factor(diabete_change_di))

# auc value
# roc(chs_pred1$diabete_change_di, chs_pred1$pred)
# ci.auc(chs_pred1$diabete_change_di, chs_pred1$pred)
# 
# vif(fit_covars)
```

## CHS model2: general risk factors + specific risk factors
```{r grf srf chs, echo=TRUE}
chs_data_model <- chs_data %>% 
  select(diabete_change_di,
         all_of(covars1),
         all_of(covars2))

fit <- glm(diabete_change_di ~.,
                 data = chs_data_model,
                 family = "binomial")

result_df2<- fit %>% tidy(., conf.int = TRUE) %>%
  mutate(model = "GRF + SRF",
         odds_ratio = exp(estimate),
         exp_conf_high = exp(conf.high),
         exp_conf_low = exp(conf.low))


chs_pred2<- chs_data %>% 
  dplyr::select(outcomes[2]) %>%
  dplyr::mutate(pred = data.frame(pred = predict(fit,
                               type = "response"))$pred) %>%
  mutate(diabete_change_di = factor(diabete_change_di))

# auc value
# roc(chs_pred2$diabete_change_di, chs_pred2$pred)
# ci.auc(chs_pred2$diabete_change_di, chs_pred2$pred)
# 
# vif(fit)
```

## CHS model3: general risk factors + selected metabolites
```{r grf met chs, echo=TRUE}
chs_data_model <- chs_data %>% 
  select(diabete_change_di,
         all_of(selected_feature_name),
         all_of(covars1)) %>%
   mutate(diabete_change_di = factor(diabete_change_di))

# logistic regression
fit <- glm(diabete_change_di~.,
           family = "binomial",
           data = chs_data_model) 

result_df3 <- fit %>% tidy(., conf.int = TRUE) %>% 
  mutate(model = "GRF + metabolites",
         odds_ratio = exp(estimate),
         exp_conf_high = exp(conf.high),
         exp_conf_low = exp(conf.low))
  
chs_pred3 <- chs_data_model %>% 
  dplyr::select(outcomes[2]) %>%
  dplyr::mutate(pred = data.frame(pred = predict(fit, type = "response"))$pred)

# roc(chs_pred3$diabete_change_di, chs_pred3$pred)
# ci.auc(chs_pred3$diabete_change_di, chs_pred3$pred)
# 
# vif(fit)
```


## CHS model4: general risk factors + specific risk factors + selected features
```{r grf srf met chs, echo=TRUE}
chs_data_model <- chs_data %>%
  select(diabete_change_di,
         all_of(selected_feature_name),
         all_of(covars1),
         all_of(covars2)) %>%
   mutate(diabete_change_di = factor(diabete_change_di))

# logistic regression
fit <- glm(diabete_change_di~.,
           family = "binomial",
           data = chs_data_model) 

result_df4 <- fit %>% tidy(., conf.int = TRUE) %>% 
  mutate(model = "GRF + SRF + metabolites",
         odds_ratio = exp(estimate),
         exp_conf_high = exp(conf.high),
         exp_conf_low = exp(conf.low))
  
# Adding annotation information to the result
model_result_annotation <- result_df4 %>% tidylog::left_join(mzrt_key_sub %>% mutate(feature = str_c("`", name ,"`")), by = c("term"= "feature")) %>% 
  select(-name, -pathway, -query_mass, -retention_time, -mass_diff,
         -matched_compound, -empirical_compound, -met_name, -path_2, term,
         met_name_first, everything())

write_csv(model_result_annotation, fs::path(dir_results,"prediction", "4_chs_model_result_with_annotation_5000.csv"))


chs_pred4 <- chs_data_model %>% 
  dplyr::select(outcomes[2]) %>%
  dplyr::mutate(pred = data.frame(pred = predict(fit, type = "response"))$pred)

# Calculate the VIF for each predictor variable in the model
(vif_df <- data.frame(vif_full_model = vif(fit)) %>% rownames_to_column("predictor"))

```

# CHS model5: selected metabolites
```{r met chs, echo=TRUE}
chs_data_model <- chs_data %>% 
  select(diabete_change_di,
         all_of(selected_feature_name)) %>%
   mutate(diabete_change_di = factor(diabete_change_di))

# logistic regression
fit <- glm(diabete_change_di~.,
           family = "binomial",
           data = chs_data_model) 

result_df5 <- fit %>% tidy(., conf.int = TRUE) %>% 
  mutate(model = "metabolites",
         odds_ratio = exp(estimate),
         exp_conf_high = exp(conf.high),
         exp_conf_low = exp(conf.low))

chs_pred5 <- chs_data_model %>% 
  dplyr::select(outcomes[2]) %>%
  dplyr::mutate(pred = data.frame(pred = predict(fit, type = "response"))$pred)

# roc(chs_pred5$diabete_change_di, chs_pred5$pred)
# ci.auc(chs_pred5$diabete_change_di, chs_pred5$pred)

```



# CHS model6: diabete_change_di ~ three random selected features 1000 times
```{r random feature chs, echo=TRUE}

foreach(i = 1:1000) %dopar% {
  # solar ----------------
  print(i)
  sample_id <- sample(1:595,3)
  
  chs_data_model <- chs_data %>% 
    select(diabete_change_di,
           features[sample_id]) %>%
     mutate(diabete_change_di = factor(diabete_change_di))
  
  # logistic regression
  fit <- glm(diabete_change_di~.,
             family = "binomial",
             data = chs_data_model) 
  
  result <- fit %>% tidy(., conf.int = TRUE) %>% 
    mutate(odds_ratio = exp(estimate),
           exp_conf_high = exp(conf.high),
           exp_conf_low = exp(conf.low))
  
  prediction <- chs_data_model %>% 
    dplyr::select(outcomes[2]) %>%
    dplyr::mutate(pred = data.frame(pred = predict(fit, type = "response"))$pred)
  
  # roc(prediction$diabete_change_di, prediction$pred)
  # auc(prediction$diabete_change_di, prediction$pred)
  
  auc_df <- data_frame(auc = auc(prediction$diabete_change_di, prediction$pred))
  
  # write_csv(auc_df,
  #           fs::path(dir_results, 
  #                    "prediction",
  #                    "4_auc_random_features.csv"),
  #           append = TRUE)
}

auc_df <- read_csv(fs::path(dir_results, 
                   "prediction",
                   "4_auc_random_features.csv"),col_names = FALSE) %>%
  rename(auc = X1)

auc_df %>% ggplot(aes(x = auc)) + 
  geom_histogram(fill = "white", color = "black") +
  labs(title = "Distribution of AUC", x = "AUC", y = "Count") +
  theme_classic()

# ggsave(filename = fs::path(dir_reports, "4_auc_distribution_random_reatures.png"), 
#        width = 10, height = 6, dpi=300)

table1::table1(~., data = auc_df)
```


# CHS model7: diabete_change_di ~ three random selected features + grf + srf 1000 times
```{r random feature and covars chs, echo=TRUE}

foreach(i = 1:1000) %dopar% {
  # solar ----------------
  print(i)
  sample_id <- sample(1:595,3)
  
  chs_data_model <- chs_data %>% 
    select(diabete_change_di,
           all_of(covars1),
           all_of(covars2),
           features[sample_id]) %>%
     mutate(diabete_change_di = factor(diabete_change_di))
  
  # logistic regression
  fit <- glm(diabete_change_di~.,
             family = "binomial",
             data = chs_data_model) 
  
  result <- fit %>% tidy(., conf.int = TRUE) %>% 
    mutate(odds_ratio = exp(estimate),
           exp_conf_high = exp(conf.high),
           exp_conf_low = exp(conf.low))
  
  prediction <- chs_data_model %>% 
    dplyr::select(outcomes[2]) %>%
    dplyr::mutate(pred = data.frame(pred = predict(fit, type = "response"))$pred)
  
  # roc(prediction$diabete_change_di, prediction$pred)
  # auc(prediction$diabete_change_di, prediction$pred)
  
  auc_df <- data_frame(auc = auc(prediction$diabete_change_di, prediction$pred))
  
  # write_csv(auc_df,
  #           fs::path(dir_results, 
  #                    "prediction",
  #                    "4_auc_random_features_covars.csv"),
  #           append = TRUE)
}

```

## AUC dataframe of all models in SOLAR and CHS
```{r aucs of models, echo=TRUE}
auc1_solar <- data.frame(auc = ci.auc(solar_pred1$diabete_change_di, solar_pred1$pred)[2], ci_low = ci.auc(solar_pred1$diabete_change_di, solar_pred1$pred)[1], ci_high = ci.auc(solar_pred1$diabete_change_di, solar_pred1$pred)[3], model = "GRF", cohort = "SOLAR", note = "General Risk Factor")


auc2_solar <- data.frame(auc = ci.auc(solar_pred2$diabete_change_di, solar_pred2$pred)[2], ci_low = ci.auc(solar_pred2$diabete_change_di, solar_pred2$pred)[1], ci_high = ci.auc(solar_pred2$diabete_change_di, solar_pred2$pred)[3], model = "GRF + SRF", cohort = "SOLAR", note = "General Risk Factor + Specific Risk Factor")


auc3_solar <- data.frame(auc = ci.auc(solar_pred3$diabete_change_di, solar_pred3$pred)[2], ci_low = ci.auc(solar_pred3$diabete_change_di, solar_pred3$pred)[1], ci_high = ci.auc(solar_pred3$diabete_change_di, solar_pred3$pred)[3], model = "GRF + metabolites", cohort = "SOLAR", note = "General Risk Factor + Lasso-selected Metabolite")

auc4_solar <- data.frame(auc = ci.auc(solar_pred4$diabete_change_di, solar_pred4$pred)[2], ci_low = ci.auc(solar_pred4$diabete_change_di, solar_pred4$pred)[1], ci_high = ci.auc(solar_pred4$diabete_change_di, solar_pred4$pred)[3], model = "GRF + SRF + metabolites", cohort = "SOLAR", note = "General Risk Factor + Specific Risk Factor + Lasso-selected Metabolite")

auc5_solar <- data.frame(auc = ci.auc(solar_pred5$diabete_change_di, solar_pred5$pred)[2], ci_low = ci.auc(solar_pred5$diabete_change_di, solar_pred5$pred)[1], ci_high = ci.auc(solar_pred5$diabete_change_di, solar_pred5$pred)[3], model = "metabolites", cohort = "SOLAR", note = "Lasso-selected Metabolite")

auc1_chs<- data.frame(auc = ci.auc(chs_pred1$diabete_change_di, chs_pred1$pred)[2], ci_low = ci.auc(chs_pred1$diabete_change_di, chs_pred1$pred)[1], ci_high = ci.auc(chs_pred1$diabete_change_di, chs_pred1$pred)[3], model = "GRF", cohort = "CHS", note = "General Risk Factor")

auc2_chs<- data.frame(auc = ci.auc(chs_pred2$diabete_change_di, chs_pred2$pred)[2], ci_low = ci.auc(chs_pred2$diabete_change_di, chs_pred2$pred)[1], ci_high = ci.auc(chs_pred2$diabete_change_di, chs_pred2$pred)[3], model = "GRF + SRF", cohort = "CHS", note = "General Risk Factor + Specific Risk Factor")

auc3_chs <- data.frame(auc = ci.auc(chs_pred3$diabete_change_di, chs_pred3$pred)[2], ci_low = ci.auc(chs_pred3$diabete_change_di, chs_pred3$pred)[1], ci_high = ci.auc(chs_pred3$diabete_change_di, chs_pred3$pred)[3], model = "GRF + metabolites", cohort = "CHS", note = "General Risk Factor + Lasso-selected Metabolite")

auc4_chs<- data.frame(auc = ci.auc(chs_pred4$diabete_change_di, chs_pred4$pred)[2], ci_low = ci.auc(chs_pred4$diabete_change_di, chs_pred4$pred)[1], ci_high = ci.auc(chs_pred4$diabete_change_di, chs_pred4$pred)[3], model = "GRF + SRF + metabolites", cohort = "CHS", note = "General Risk Factor + Specific Risk Factor + Lasso-selected Metabolite")

auc5_chs <- data.frame(auc = ci.auc(chs_pred5$diabete_change_di, chs_pred5$pred)[2], ci_low = ci.auc(chs_pred5$diabete_change_di, chs_pred5$pred)[1], ci_high = ci.auc(chs_pred5$diabete_change_di, chs_pred5$pred)[3], model = "metabolites", cohort = "CHS", note = "Lasso-selected Metabolite")

auc <- auc1_solar %>% 
  bind_rows(auc2_solar) %>%
  bind_rows(auc3_solar) %>%
  bind_rows(auc4_solar) %>%
  bind_rows(auc5_solar) %>%
  bind_rows(auc1_chs) %>%
  bind_rows(auc2_chs) %>%
  bind_rows(auc3_chs) %>%
  bind_rows(auc4_chs) %>%
  bind_rows(auc5_chs) %>%
  mutate(cohort = factor(cohort, levels = c("SOLAR", "CHS")))

# write_csv(auc, 
#           fs::path(dir_results, 
#                    "prediction",
#                    "4_auc_df.csv"))

auc
```

